Generalized Tractability

### 2023-03-30 GeneralizedTractability_230310.tex
Fred Office computer updates

Next we show that if $\vp^* \in \widetilde{\cp}_{\textup{strct}}$, then $\vp^* \in \cp_{\textup{strct}}$.  If $\vp^* \in \widetilde{\cp}_{\textup{strct}}$, then \eqref{eq:strong_tractiff3}  holds for all $\vp \in (\vp^*,\vinfty)$.  By the argument to prove the sufficient condition above, it follows that \eqref{eq:strong_tract_def} must also hold for all $\vp \in (\vp^*,\vinfty)$. Thus, $\vp^* \in \cp}_{\textup{strct}}$.

@@

\begin{theorem}\label{thm_main_tract2}
A problem is $T$-tractable iff there exist $\vp\ge \vzero$ and a constant $L(\vp) > 0$ such that
\begin{equation} \label{eq:tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L(\vp)\, T(0,d,\vp) \rceil}^\infty \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:tractiff4} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{trct}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$.  Then $\cp_{\textup{trct}} = \widetilde{\cp}_{\textup{trct}}$, and the set of optimal strong tractability parameters is
\[
\cp_{\textup{opt}} =
\{\vp^* \in \widetilde{\cp}_{\textup{trct}} :  \vp* \notin [\vp^\dagger,\boldsymbol{\infty}) \ \forall \vp^\dagger \in  \widetilde{\cp}_{\textup{trct}} \setminus \{\vp^*\} \}.
\]\end{theorem}

@@

Let $J(\vp,d): = \lceil L(\vp)\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & = 1 + T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}} \Biggr ] \\
       & \le 1+ T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]
        \\
       & \le 1+ T(\varepsilon^{-1},d,\vp)\left[L(\vp)\, T(0,d,\vp) + S_{\vp}\right] \qquad \text{since } T(\lambda_{\max}^{-1},d,\vp) \ge 1\\
       & \le 1+L(\vp)\,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le 1+L(\vp)\,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le 1+L(\vp)\,K_{\vp,2} T(\varepsilon^{-1},d,2\vp)  + S_{\vp}\, T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le C_{2}(2\vp) \,T(\varepsilon^{-1},d,2\vp)
\end{align*}
for some suitably chosen $C_{2}(2\vp)>0$. It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}. \\

@@

For $\varepsilon$ taking on all positive values, $j(\varepsilon,d,\vp)$ takes on all values greater than or equal to $j(\infty,d,\vp)$, so
\[
j\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{j,d}^{-1},d,\vp)+1, \qquad \forall j \ge j(\infty,d,\vp).
\]
This implies via our technical assumption \eqref{eq:ptauassume} that
\begin{equation*}
	K_{\vp,\tau}\,T (\lambda_{j,d}^{-1},d,\tau \vp) \ge
	[T(\lambda_{j,d}^{-1},d, \vp)]^\tau
	\ge
	\left[\frac{(j-1)}{C_{\vp}}\right]^\tau \qquad \forall j \ge j(\infty,d,\vp)
\end{equation*}
Summing both sides of the inequality yields and noting that $L(\vp) \le L(\tau\vp)$, it follows that
\peter{(Peter: Why does $L(\cdot)$ ocur here?)} \fred{Because I got lost a bit.}

@@

Let $T: (0,\infty) \times \naturals \times [0,\infty) \times [0,\infty)\Rightarrow [1,\infty)$
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= (\max\{\varepsilon^{-1},1\})^p\, d^q.
\]
Then Theorem \ref{thm_main_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L(\vp) d^{q_1}}^\infty (\min\{\lambda_{i,d},1\})^p d^{-q} < \infty.
\]
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08}

>>>>

Next we show that if $\vp^* \in \widetilde{\cp}_{\textup{strct}}$, then $\vp^* \in \cp_{\textup{strct}}$.  If $\vp^* \in \widetilde{\cp}_{\textup{strct}}$, then \eqref{eq:strong_tractiff3}  holds for all $\vp \in (\vp^*,\vinfty)$.  By the argument to prove the sufficient condition above, it follows that \eqref{eq:strong_tract_def} must also hold for all $\vp \in (\vp^*,\vinfty)$. Thus, $\vp^* \in \cp_{\textup{strct}}$.

@@

The argument proving equivalent conditions for tractability is similar to, but somewhat more involved than the proof of Theorem \ref{thm_main_strong_tract2}.  Moreover, the specification of the set of optimal parameters is not as precise as in Theorem \ref{thm_main_strong_tract2}.

\begin{theorem}\label{thm_main_tract2}
A problem is $T$-tractable iff there exist $\vp\ge \vzero$ and a constant $L_{\vp} > 0$ such that
\begin{equation} \label{eq:tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^\infty \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:tractiff4} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{trct}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$, and   $\widehat{\cp}_{\textup{trct}} : = \{\vp^* : 2\vp \in \widetilde{\cp}_{\textup{trct}}\}$.  Then $\cp_{\textup{trct}} = \widetilde{\cp}_{\textup{trct}}$, and the set of optimal strong tractability parameters is
\[
\cp_{\textup{opt}} =
\{\vp^* \in \widetilde{\cp}_{\textup{trct}} :  \vp^* \notin [\vp^\dagger,\boldsymbol{\infty}) \ \forall \vp^\dagger \in  \widetilde{\cp}_{\textup{trct}} \setminus \{\vp^*\} \}.
\]\end{theorem}

@@

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & = 1 + T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}} \Biggr ] \\
       & \le 1+ T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]
        \\
       & \le 1+ T(\varepsilon^{-1},d,\vp)\left[L_{\vp}\, T(0,d,\vp) + S_{\vp}\right] \qquad \text{since } T(\lambda_{\max}^{-1},d,\vp) \ge 1\\
       & \le 1+L_{\vp}\,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le 1+L_{\vp}\,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le 1+L_{\vp}\,K_{\vp,2} T(\varepsilon^{-1},d,2\vp)  + S_{\vp}\, T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le C_{2}(2\vp) \,T(\varepsilon^{-1},d,2\vp)
\end{align*}
for some suitably chosen $C_{2}(2\vp)>0$. It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}. \\

@@

For $\varepsilon$ taking on all positive values, $j(\varepsilon,d,\vp)$ takes on all values greater than or equal to $j(\infty,d,\vp)$, so
\[
j\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{j,d}^{-1},d,\vp)+1, \qquad \forall j \ge j(\infty,d,\vp).
\]
This implies via our technical assumption \eqref{eq:ptauassume} that
\begin{equation*}
	K_{\vp,\tau}\,T (\lambda_{j,d}^{-1},d,\tau \vp) \ge
	[T(\lambda_{j,d}^{-1},d, \vp)]^\tau
	\ge
	\left[\frac{(j-1)}{C_{\vp}}\right]^\tau \qquad \forall j \ge j(\infty,d,\vp)
\end{equation*}
Summing both sides of the inequality yields and noting that $L_{\vp} \le L(\tau\vp)$, it follows that
\peter{(Peter: Why does $L(\cdot)$ ocur here?)} \fred{Because I got lost a bit.}

@@

Let $T: (0,\infty) \times \naturals \times [0,\infty) \times [0,\infty)\Rightarrow [1,\infty)$
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= (\max\{\varepsilon^{-1},1\})^p\, d^q.
\]
Then Theorem \ref{thm_main_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L_{\vp} d^{q_1}}^\infty (\min\{\lambda_{i,d},1\})^p d^{-q} < \infty.
\]
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08}
